---
title: "Fall Cohort 2 analysis - 17 dpi - Cohort ~3~"
author: "Brandon Roy "
date: "1/5/23"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}

library(tidyverse)
library(ggpubr)
library(rstatix)
library(readxl)
library(gplots)
library(corrplot)

```


```{r}
Root_Summary_2 <- read_excel("Fall Cohort2_Rhizovision data.xlsx")
metadata_2 <- read_excel("Fall cohort2_measurements.xlsx")
Root_Summary_2 <- na.omit(Root_Summary_2)
Root_Summary_2 <- as.data.frame(Root_Summary_2)
# variable assignment
treatment2 <- Root_Summary_2$Treatment
tips2 <- Root_Summary_2$`number of root tips`
tot_length2 <- Root_Summary_2$`total root length (mm)`
branch_points2 <- Root_Summary_2$`number of branch points`
branch_freq2 <- Root_Summary_2$`branching frequency (mm^-1)`
network2 <- Root_Summary_2$`network area (mm^2)`
diameter2 <- Root_Summary_2$`average diameter (mm)`
surface_area2 <- Root_Summary_2$`surface area (mm^2)`
volume2 <- Root_Summary_2$`volume (mm^3)`
height2 <- metadata_2

height2 <- metadata_2$`Height (mm)`
Category2 <- metadata_2$treatment
Biomass_Final2 <- metadata_2$`total dry mass (paper bag + tissue)`


summ <- Root_Summary_2 %>%
  group_by(treatment) %>%
  summarize(n=n(), 'number of root tips' = mean(`number of root tips`))
summ2 <- Root_Summary_2 %>%
  group_by(treatment) %>%
  summarize(n=n(), 'average diameter' = mean(`average diameter (mm)`))
summ3 <- Root_Summary_2 %>%
  group_by(treatment) %>%
  summarize(n=n(), 'branching frequency' = mean(`branching frequency`))
#### Plot the data - reponse variables with repect to treatment type

    # The five treatments are mock inoculation, GFLV GHu wild type strain, GFLV F13 wild type strain, 
    # GFLV GHu 1EPol K802G strain, and GFLV F13 1EPol G802K N804S strain.
Root_Summary_2 %>%
  group_by(treatment) %>%
  summarise(
    count = n(),
    `height (mm)` = mean(`height (mm)`),
    `dry root biomass (g)` = mean(`dry root biomass (g)`),
    `dry shoot biomass (g)` = mean(`dry shoot biomass (g)`),
    `root to shoot ratio` = mean(`root to shoot ratio`),
    `number of root tips` = mean(`number of root tips`),
    `number of branch points` = mean(`number of branch points`),
    `total root length (mm)` = mean(`total root length (mm)`),
    `branching frequency` = mean(`branching frequency`),
    `network area (mm^2)` = mean(`network area (mm^2)`),
    `average diameter (mm)` = mean(`average diameter (mm)`),
    `median diameter (mm)` = mean(`median diameter (mm)`),
    `perimeter (mm)` = mean(`perimeter (mm)`),
    `volume (mm^3)` = mean(`volume (mm^3)`),
    `surface area (mm^2)` = mean(`surface area (mm^2)`)
  )

```

```{r}
#Height
metadata_2$`plant height (mm)`
Height_plot <- ggboxplot(metadata_2, x = "treatment", y= "Height (mm)", fill = "treatment", palette = colorlist, xlab = FALSE, width = 0.5, ylab = "Height (mm)")+
  stat_compare_means(label = 'p.signif', method ='wilcox.test', ref.group = "control", hide.ns = TRUE)
#Dry Biomass
Dry_shoot_plot <- ggboxplot(metadata_2, x = "treatment", y= "dry shoot biomass (g)", fill = "treatment", palette = colorlist, xlab = FALSE, width = 0.5, ylab = "Dry biomass (g)")+
  stat_compare_means(label = 'p.signif', method ='wilcox.test', ref.group = "control", hide.ns = TRUE)
Dry_root_plot <- ggboxplot(metadata_2, x = "treatment", y= "dry root biomass (g)", fill = "treatment", palette = colorlist, xlab = FALSE, width = 0.5, ylab = "Dry biomass (g)")+
  stat_compare_means(label = 'p.signif', method ='wilcox.test', ref.group = "control", hide.ns = TRUE)
#Volume
Vol_plot <- ggboxplot(Root_Summary_2, x = "treatment", y= "volume (mm^3)", fill = "treatment", palette = colorlist, xlab = FALSE, width = 0.5, ylab = "Volume (mm^3)")+
  stat_compare_means(label = 'p.signif', method ='wilcox.test', ref.group = "control", hide.ns = TRUE)


multi_plot2 <- ggarrange(Height_plot + rremove("x.text"), Dry_shoot_plot + rremove("x.text"), Dry_root_plot + rremove("x.text"), Vol_plot + rremove("x.text"), common.legend = TRUE, legend = 'bottom',labels = 'AUTO')
multi_plot2

ggviolin(Root_Summary_2, x = "treatment", y= "total root length (mm)")
ggboxplot(Root_Summary_2, x = "treatment", y= "number of branch points")
ggboxplot(Root_Summary_2, x = "treatment", y= "network area (mm^2)")
ggboxplot(Root_Summary_2, x = "treatment", y= "average diameter (mm)")
```


```{r}
plotmeans(Root_Summary_2$`number of root tips` ~ Root_Summary_2$treatment, xlab="treatment", ylab="Number of Tips", main="Root Hairs with respect to infection by GFLV", data = Root_Summary_2)
```

```{r}
#Root tips 
ggboxplot(na.omit(Root_Summary_2), x = "treatment", y= "number of root tips", fill = "treatment", palette =colorlist, labels = FALSE, xlab = FALSE, width = 0.5) +
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE, text.size = 20) + 
  theme(aspect.ratio = 1)
```
```{r}
#Total Root Length
rtlen <- ggboxplot(Root_Summary_2, x = "treatment", y= "total root length (mm)", fill = "treatment", palette = colorlist, xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='wilcox.test', ref.group = "Control", hide.ns = TRUE) + theme(aspect.ratio = 1)
#Number of Branch Points
bp <- ggboxplot(Root_Summary_2, x = "treatment", y= "number of branch points", fill = "treatment", palette = colorlist, xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)+ theme(aspect.ratio = 1)
#Average Diameter
dia <- ggboxplot(Root_Summary_2, x = "treatment", y= "average diameter (mm)", fill = "treatment", palette = colorlist, xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)+ theme(aspect.ratio = 1)

# create multi panel plot
multi_plot <- ggarrange(rtip + rremove("x.text"), rtlen + rremove("x.text"),  bp + rremove("x.text"),dia + rremove("x.text"), common.legend = TRUE, legend = 'top',labels = 'AUTO', rows=4, cols=1, font.label = list(size=20)) + theme(aspect.ratio =0.8)
multi_plot
ggarrange(rtip + rremove("x.text"), rtlen + rremove("x.text"), bp, dia, common.legend = TRUE, labels = 'AUTO')
```

```{r}

#Height
he1 <- ggboxplot(metadata_2, x = "treatment", y= "Height (mm)", fill = "treatment", palette = colorlist, xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='anova', ref.group = "control", hide.ns = TRUE)
he1
#Volume
rtvol1 <- ggboxplot(Root_Summary_2, x = "treatment", y= "Volume", fill = "Treatment", palette = colorlist, xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "control", hide.ns = TRUE)
#Network Area
rtnet1 <- ggboxplot(Root_Summary_2, x = "treatment", y= "Network area (mm^2)", fill = "Treatment", palette = colorlist, xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "control", hide.ns = TRUE)
#Surface Area
rtsa1 <- ggboxplot(Root_Summary_2, x = "treatment", y= "Surface area", fill = "Treatment", palette = colorlist, xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "control", hide.ns = TRUE)

#combine everything together for a 4-panel figure
metaplot1 <- ggarrange(he + rremove("x.text"), rtvol + rremove("x.text"), rtnet + rremove("x.text"), rtsa + rremove("x.text"), Total_mass_plot + rremove("x.text"), common.legend = TRUE, legend = 'bottom',labels = 'AUTO')
metaplot1

```


```{r}
GHu_Summary_3 <- subset(Root_Summary_2, treatment == 'wildtype GHu')
lm_symptoms3 <- lm(`number of root tips` ~ `days showing symptoms`, data=GHu_Summary3)
summary(lm_symptoms3)
plot(lm_symptoms3)

lm_symptoms3 <- lm(`average diameter (mm)` ~ `days showing symptoms`, data=GHu_Summary3)

ggplot(GHu_Summary3, aes(x=`days showing symptoms`, y=`average diameter (mm)`)) + 
  geom_point() +
  geom_smooth(method=lm, color='turquoise3')+ 
  theme(text = element_text(size = 20), axis.text = element_text(size = 16), aspect.ratio = 0.4)
ggplot(GHu_Summary_3, aes(x=`days showing symptoms`, y=`number of root tips`)) + 
  geom_point() +
  geom_smooth(method=lm, color='red')+ 
  theme(text = element_text(size = 20), axis.text = element_text(size = 16), aspect.ratio = 0.4)

GHu_Summary3 <- subset(GHu_Summary_3, `days showing symptoms` > 1)
ggplot(GHu_Summary3, aes(x=`days showing symptoms`, y=`median diameter (mm)`)) + 
  geom_point() +
  geom_smooth(method=lm)
ggplot(GHu_Summary3, aes(x=`days showing symptoms`, y=`number of root tips`)) + 
  geom_point() +
  geom_smooth(method=lm)

# mixed model construction

# linear modeling approach utilizing the first or last day symptoms are shown (and both)

mixed_lm_tips4 <- lm(`number of root tips` ~ `first symptomatic` + `last symptomatic` + `first symptomatic`:`last symptomatic`, data=GHu_Summary3)
summary(mixed_lm_tips4)
plot(mixed_lm_tips4)
ggplot(GHu_Summary3, aes(x=`days showing symptoms`, y=mixed_lm_tips4$fitted.values)) + 
  geom_point() +
  geom_smooth(method=lm, color ='turquoise4')
mixed_lm_tips4 <- lm(`surface area (mm^2)` ~ `first symptomatic` + `last symptomatic` + `first symptomatic`:`last symptomatic`, data=GHu_Summary3)
summary(mixed_lm_tips4)
plot(mixed_lm_tips4)
ggplot(GHu_Summary3, aes(x=`days showing symptoms`, y=mixed_lm_tips4$fitted.values)) + 
  geom_point() +
  geom_smooth(method=lm, color ='turquoise4')


library(car)
avPlots(lm_tips)

lm_lastdia <- lm(`average diameter (mm)` ~ `last symptomatic`, data=GHu_Summary3)
lm_firstdia <- lm(`average diameter (mm)` ~ `first symptomatic`, data=GHu_Summary3)
summary(lm_lastdia)
summary(lm_firstdia)

ggplot(GHu_Summary3, aes(x=`last symptomatic`, y=`average diameter (mm)`)) + 
  geom_point() +
  geom_smooth(method=lm, color ='red')
ggplot(GHu_Summary3, aes(x=`first symptomatic`, y=`average diameter (mm)`)) + 
  geom_point() +
  geom_smooth(method=lm, color ='red')
ggplot(GHu_Summary3, aes(x=`last symptomatic`, y=`number of root tips`)) + 
  geom_point() +
  geom_smooth(method=lm, color ='red')
ggplot(GHu_Summary3, aes(x=`first symptomatic`, y=`number of root tips`)) + 
  geom_point() +
  geom_smooth(method=lm, color ='red')
```

```{r}
#correlation analysis of late cohort
root_sum_matrix2 <- Root_Summary_2[c(3,6:23)]

root_cor2 <- cor(root_sum_matrix2, use='complete.obs')
res1 = cor.mtest(root_cor2, conf.level=0.99)
corrplot(root_cor2, type = 'full', order='hclust', hclust.method='average', tl.col='black', tl.srt=45, addrect=6, method="circle", p.mat = res1$p, sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.9, insig = 'label_sig', pch.col = 'grey20')
corrplot(root_cor2, type = 'full', order='hclust', hclust.method='average', tl.col='black', tl.srt=45, addrect=6, method="circle", p.mat = res1$p, sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.9, insig = 'label_sig', pch.col = 'grey20')


## add significant level stars and cluster rectangles
corrplot(M, p.mat = testRes$p, tl.pos = 'd', order = 'hclust', addrect = 2,
         insig = 'label_sig', sig.level = c(0.001, 0.01, 0.05),
         pch.cex = 0.9, pch.col = 'grey20')
corrplot.mixed(root_cor2, order = 'hclust', ) %>%
  corrRect(name = c('scan side', 'median diameter (mm)', 'root to shoot ratio'))
```

```{r}
#### THIS BLOCK OF CODE NEEDS MODIFIED #####
typeof(Root_Summary_2)
res_tip22_2 <- kruskal.test(Root_Summary_2, `number of root tips` ~ treatment)
pwc_tip22_2 <- Root_Summary_2 %>%
  dunn_test(`number of root tips` ~ `treatment`, p.adjust.method = "bonferroni")
pwc_tip22_2

ggviolin(Root_Summary_2, x = "treatment", y= "number of root tips", fill = "treatment", palette =colorlist, xlab = FALSE, alpha = 1, remove = NA)+  
  geom_boxplot(width = .2, alpha = 0.9, fatten = NULL, show.legend = TRUE, legend='right') +
  stat_pvalue_manual(pwc_tip22_2, hide.ns=T, y.position =760, step.increase=0.1) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.position = 'right') 
#  labs(subtitle = get_test_label(res_tip22_2), detailed = T, caption = get_pwc_label(pwc_tip22_2)) 

res_dia222 <- kruskal.test(Root_Summary_2, `average diameter (mm)` ~ treatment)
pwc_dia222 <- Root_Summary_2 %>%
  dunn_test(`average diameter (mm)` ~ `treatment`, p.adjust.method = "bonferroni")
pwc_dia222

ggviolin(Root_Summary_2, x = "treatment", y= "average diameter (mm)", fill = "treatment", palette = colorlist, xlab = FALSE, alpha = 1, remove=NA)+  
  geom_boxplot(width = .2, alpha = 0.9, fatten = NULL, show.legend = TRUE, legend='right') +
  stat_pvalue_manual(pwc_dia222, hide.ns=T, y.position = 1.35, step.increase=0.1) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.position = 'right')
#  labs(subtitle = get_test_label(res_dia222), detailed = T, caption = get_pwc_label(pwc_dia222)) 

res_br222 <- kruskal.test(Root_Summary_2, `number of branch points` ~ treatment)
pwc_br222 <- Root_Summary_2 %>%
  dunn_test(`number of branch points` ~ `treatment`, p.adjust.method = "bonferroni")
pwc_br222

ggviolin(Root_Summary_2, x = "treatment", y="number of branch points", fill = "treatment", palette = colorlist, xlab = FALSE, alpha = 1, remove=NA)+  
  geom_boxplot(width = .2, alpha = 0.9, fatten = NULL, show.legend = TRUE, legend='right') +
  stat_pvalue_manual(pwc_br222, hide.ns=T, y.position = .7, step.increase=0.1) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.position = 'right')
#  labs(subtitle = get_test_label(res_dia222), detailed = T, caption = get_pwc_label(pwc_dia222)) 
```

```{r}
ggviolin(Root_Summary_2, x = "treatment", y= "number of root tips", fill = "treatment", palette =colorlist, remove=NA) + 
  geom_boxplot(width = .2, alpha = .7, fatten = NULL)+
  stat_compare_means(aes(label = ..p.signif..), method="t.test",ref.group = "control", label.y = c(640,750,600,450,425), hide.ns = T) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  stat_compare_means(method="anova",label.y=775) +
  theme(aspect.ratio = 1) +
  theme(axis.text.x = element_blank(), legend.position = 'none')

```

```{r}
ggviolin(Root_Summary_2, x = "treatment", y="average diameter (mm)", fill = "treatment", palette =colorlist2, remove=NA) + 
  geom_boxplot(width = .2, alpha = .7, fatten = NULL)+
  stat_compare_means(aes(label = ..p.signif..), method="t.test",ref.group = "control", label.y = c(1,1.3,1,1,1.3),  hide.ns = T) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  stat_compare_means(method="anova",label.y=1.4) +
  theme(aspect.ratio = 1) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

ggviolin(Root_Summary_2, x = "treatment", y= "number of branch points", fill = "treatment", palette =colorlist2, remove=NA) + 
  geom_boxplot(width = .2, alpha = .7, fatten = NULL)+
  stat_compare_means(aes(label = ..p.signif..), method="t.test",ref.group = "control", hide.ns = T) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  stat_compare_means(method="anova",label.y=1900) +
  theme(aspect.ratio = 1) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

ggviolin(Root_Summary_2, x = "treatment", y= "volume (mm^3)", fill = "treatment", palette =colorlist3, remove=NA) + 
  geom_boxplot(width = .2, alpha = .7, fatten = NULL)+
  stat_compare_means(aes(label = ..p.signif..), method="t.test",ref.group = "control", hide.ns = T) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  stat_compare_means(method="anova",label.y=3300) +
  theme(aspect.ratio = 1) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

ggviolin(Root_Summary_2, x = "treatment", y= "total root length (mm)", fill = "treatment", palette =colorlist2, remove=NA) + 
  geom_boxplot(width = .2, alpha = .7, fatten = NULL)+
  stat_compare_means(aes(label = ..p.signif..), method="t.test",ref.group = "Control", hide.ns = T) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  stat_compare_means(method="anova",label.y=3000) +
  theme(aspect.ratio = 1) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

ggviolin(Root_Summary_2, x = "treatment", y= "surface area (mm^2)", fill = "treatment", palette =colorlist2, remove=NA) + 
  geom_boxplot(width = .2, alpha = .7, fatten = NULL)+
  stat_compare_means(aes(label = ..p.signif..), method="t.test",ref.group = "Control", hide.ns = T) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  stat_compare_means(method="anova",label.y=7500) +
  theme(aspect.ratio = 1) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r}

# height
mod <- lm(`height (mm)` ~ treatment, data = Root_Summary_2)
mod_means_contr <- emmeans::emmeans(object = mod,
                                    pairwise ~ "treatment",
                                    adjust = "tukey")
mod_means <- multcomp::cld(object = mod_means_contr$emmeans,
                           Letters = letters)
#number of root tips
mod <- lm(`number of root tips` ~ treatment, data = Root_Summary_2)
mod_means_contr <- emmeans::emmeans(object = mod,
                                    pairwise ~ "treatment",
                                    adjust = "tukey")
multcomp::cld(object = mod_means_contr$emmeans,
                           Letters = letters)
# average diameter
mod <- lm(`average diameter (mm)` ~ treatment, data = Root_Summary_2)
mod_means_contr <- emmeans::emmeans(object = mod,
                                    pairwise ~ "treatment",
                                    adjust = "tukey")
mod_means <- multcomp::cld(object = mod_means_contr$emmeans,
                           Letters = letters)
# number of branch points
mod <- lm(`number of branch points` ~ treatment, data = Root_Summary_2)
mod_means_contr <- emmeans::emmeans(object = mod,
                                    pairwise ~ "treatment",
                                    adjust = "sidak")
mod_means <- multcomp::cld(object = mod_means_contr$emmeans,
                           Letters = letters)

mod <- lm(`average diameter (mm)` ~ treatment, data = Root_Summary_2)
mod_means_contr <- emmeans::emmeans(object = mod,
                                    pairwise ~ "treatment",
                                    adjust = "tukey")
mod_means <- multcomp::cld(object = mod_means_contr$emmeans,
                           Letters = letters)


library(ggplot2)

ggplot(data = mod_means,
       aes(x = treatment, y = emmean)) +
  geom_errorbar(aes(ymin = lower.CL, 
                    ymax = upper.CL), 
                width = 0.2) +
  geom_point() +
  geom_text(aes(label = gsub(" ", "", .group)),
            position = position_nudge(x = 0.2)) +
  labs(caption = "Means followed by a common letter are\nnot significantly different according to the Tukey-test")




```

#Root tips
ggboxplot(Root_Summary, x = "Treatment", y= "Root tips", fill = "Treatment", palette = "jco", xlab = FALSE, width = 0.5)+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)
#Total Root Length
ggboxplot(Root_Summary, x = "Treatment", y= "Total root length (mm)", color = "Treatment", xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)
#Number of Branch Points
ggboxplot(Root_Summary, x = "Treatment", y= "Number of branch points", color = "Treatment", xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)
#Average Diameter
ggboxplot(Root_Summary, x = "Treatment", y= "Average diameter (mm)", color = "Treatment", xlab = FALSE, width = 0.5,)+
  stat_compare_means(label = 'p.signif', method ='wilcox.test', ref.group = "Control", hide.ns = TRUE)
