---
title: "Winter Cohort 1"
author: "Brandon Roy"
date: "10.24.22"
output:
  word_document: default
  html_document: default
---

# Description of dataset
# Harvested 47 days after sowing, 16 days after inoculation
# Treatments        # No. Inoculated      # Infected
# Control               15                    0
# F13 WT                15                    5
# F13 Mutant            21                    11
# GHU WT                15                    13
# GHU Mutant            20                    17

# 25 individuals uninfected but inoculated

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(readxl)
library(gplots)
library(corrplot)
library(multcompView)
library(rcompanion)
library(DescTools)
library(FSA)
library(data.table)
colorlist <- c('#8d96a3','#00798c','light blue','#d1495b','orange')
```


```{r}
Root_Summary_1 <- read_excel("Winter Cohort 1 Rhizovision.xlsx", sheet=1)

# variable assignment
Root_Summary_1$Treatment = as.factor(Root_Summary_1$treatment)

Root_Summary_1 <- Root_Summary_1 %>% mutate(Treatment = fct_relevel(treatment, "mock control", "wildtype F13", "mutant F13 1E G802K", "wildtype GHu", "mutant GHu 1E K802G"))  
ggboxplot(Root_Summary_1, x = "treatment", y= "number of root tips")+
  stat_compare_means(method = 'anova')

# The six treatments are mock inoculation, uninfected, GFLV GHu wild type strain, GFLV GHu Mutant of 2404-2424 F13 cross,     
# GFLV F13 wild type strain, and GFLV F13 Mutant of 2404-2424 GHu cross strain.

```

```{r}
#Height
Height_plot <- ggboxplot(Root_Summary_1, x = "Treatment", y= "Height (mm)", fill = "Treatment", palette = "jco", xlab = FALSE, width = 0.5, ylab = "Height (mm)")+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)
#Volume
Vol_plot <- ggboxplot(Root_Summary_1, x = "Treatment", y= "volume (mm^3)", fill = "Treatment", palette = "jco", xlab = FALSE, width = 0.5, ylab = "Volume (mm^3)")+
  stat_compare_means(label = 'p.signif', method ='wilcox.test', ref.group = "Control", hide.ns = TRUE)

Height_plot

Vol_plot


######### WHAT CAN WE CONCLUDE? #############


ggviolin(Root_Summary_1, x = "Treatment", y= "total root length (mm)")
ggboxplot(Root_Summary_1, x = "Treatment", y= "number of branch points")
ggboxplot(Root_Summary_1, x = "Treatment", y= "network area (mm^2)")
ggboxplot(Root_Summary_1, x = "Treatment", y= "average diameter (mm)")
###

tipmodel1 <- aov(tips1 ~ Treatment, data = Root_Summary_1)
tipmodel1

aov(tipmodel1)
anova(tipmodel1)
plot(tipmodel1)


ggdensity(Root_Summary_1$`number of root tips`)
ggqqplot(Root_Summary_1$`number of root tips`)
shapiro.test(Root_Summary_1$`number of root tips`) # not a normal distribution for root tips
ggdensity(Root_Summary_1, x = "number of root tips", y= '..count..', facet.by = "Treatment")
```

```{r}
plotmeans(tips1 ~ treatment1, xlab="treatment", ylab="Number of Tips",
main="Root Hairs with respect to infection by GFLV")

Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('number of root tips', type = "mean_se")

Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('number of branch points', type = "mean_se")
Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('total root length (mm)', type = "mean_se")
Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('branching frequency (mm^-1)', type = "mean_se")
Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('network area (mm^2)', type = "mean_se")
Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('average diameter (mm)', type = "mean_se")
Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('median diameter (mm)', type = "mean_se")
Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('perimeter (mm)', type = "mean_se")
Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('volume (mm^3)', type = "mean_se")
Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('surface area (mm^2)', type = "mean_se")
Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('height (mm)', type = "mean_se")
Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('root dry biomass (g)', type = "mean_se")
Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('shoot dry biomass (g)', type = "mean_se")
Root_Summary_1 %>%
  group_by(treatment) %>%
  get_summary_stats('root to shoot ratio', type = "mean_se")
```

```{r, Statistics}

# Root tip stats 
res_tip <- Root_Summary_3 %>% kruskal_test(`number of root tips` ~ `treatment`) # reject
pwc_tip <- Root_Summary_1 %>%
  dunn_test(`number of root tips` ~ `Treatment`, p.adjust.method = "bonferroni")
pwc_tip

DT = dunnTest(`number of root tips` ~ Treatment, data = Root_Summary_1, method = 'bonferroni')
PT = DT$res
tiplabel = cldList(P.adj ~ Comparison, data = PT, threshold = 0.05)
tiplabel1 <- tiplabel %>% arrange(c("mock control", "wildtype F13", "mutant F13 1E G802K", "wildtype GHu", "mutant GHu 1E K802G"))

# Diameter stats
res_dia <- Root_Summary_1 %>% kruskal_test(`average diameter (mm)` ~ Treatment)
pwc_dia <- Root_Summary_1 %>%
  dunn_test(`average diameter (mm)` ~ `Treatment`, p.adjust.method = "bonferroni")
pwc_dia

DDT = dunnTest(`average diameter (mm)` ~ Treatment, data = Root_Summary_1, method = 'bonferroni')
PDT = DDT$res
dilabel = cldList(P.adj ~ Comparison, data = PDT, threshold = 0.05)
dilabel1 <- dilabel %>% arrange(c("mock control", "wildtype F13", "mutant F13 1E G802K", "wildtype GHu", "mutant GHu 1E K802G"))

# Branching frequency stats
res_nb <- Root_Summary_1 %>% kruskal_test(`branching frequency (mm^-1)` ~ Treatment)
pwc_nb <- Root_Summary_1 %>%
  dunn_test(`branching frequency (mm^-1)` ~ `Treatment`, p.adjust.method = "bonferroni")
pwc_nb

DBT = dunnTest(`branching frequency (mm^-1)` ~ Treatment, data = Root_Summary_1, method = 'bonferroni')
PBT = DBT$res
brlabel = cldList(P.adj ~ Comparison, data = PBT, threshold = 0.05)
brlabel1 <- brlabel %>% arrange(c("mock control", "wildtype F13", "mutant F13 1E G802K", "wildtype GHu", "mutant GHu 1E K802G"))
```



```{r}
#Root tips 
ggboxplot(Root_Summary_1, x = "Treatment", y= "number of root tips", fill = "Treatment", palette =colorlist2, labels = FALSE, xlab = FALSE, width = 0.5) +
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE, text.size = 20) + 
  theme(aspect.ratio = 1)

ggviolin(Root_Summary_1, x = "Treatment", y= "number of root tips", fill = "Treatment", palette =colorlist, xlab = FALSE, alpha = 1)+  
  geom_boxplot(width = .2, alpha = 0.9, fatten = NULL, show.legend = TRUE, legend='right') +
  stat_pvalue_manual(pwc_tip, hide.ns=T, y.position =500, step.increase=0.1) +
  labs(subtitle = get_test_label(res_tip), detailed = T, caption = get_pwc_label(pwc_tip)) + 
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.position = 'right')

p <- ggviolin(Root_Summary_1, x = "Treatment", y= "number of root tips", fill = "Treatment", palette =colorlist, xlab = FALSE, alpha = 1)+  
  geom_boxplot(width = .2, alpha = 0.9, fatten = NULL, show.legend = TRUE, legend='right') +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.position = 'right')
p + annotate("text", x = 1:6, y= 600, label = tiplabel1$Letter)

nrow(Root_Summary_1)

Root_Summary_1$Treatment

```
```{r}
# Diameter plots with stat labels
ggviolin(Root_Summary_1, x = "Treatment", y= "average diameter (mm)", fill = "Treatment", palette =colorlist, xlab = FALSE, alpha = 1)+  
  geom_boxplot(width = .2, alpha = 0.9, fatten = NULL, show.legend = TRUE, legend='right') +
  stat_pvalue_manual(pwc_dia, hide.ns=T, y.position = 1.5, step.increase=0.1) +
  labs(subtitle = get_test_label(res_dia), detailed = T, caption = get_pwc_label(pwc_dia)) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.position = 'right')

d <- ggviolin(Root_Summary_1, x = "Treatment", y= "average diameter (mm)", fill = "Treatment", palette =colorlist, xlab = FALSE, alpha = 1)+  
  geom_boxplot(width = .2, alpha = 0.9, fatten = NULL, show.legend = TRUE, legend='right') +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.position = 'right')
d + annotate("text", x = 1:5, y= 1.4, label = dilabel1$Letter)
```


```{r}
ggviolin(Root_Summary_1, x = "Treatment", y= "branching frequency (mm^-1)", fill = "Treatment", palette =colorlist, xlab = FALSE, alpha = 1)+  
  #geom_boxplot(width = .2, alpha = 0.9, fatten = NULL, show.legend = TRUE, legend='right') +
  stat_pvalue_manual(pwc_nb, hide.ns=T, y.position = 1.2, step.increase=0.1) +
  labs(subtitle = get_test_label(res_nb), detailed = T, caption = get_pwc_label(pwc_nb)) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.position = 'right')

d <- ggviolin(Root_Summary_1, x = "Treatment", y= "branching frequency (mm^-1)", fill = "Treatment", palette =colorlist, xlab = FALSE, alpha = 1)+  
  geom_boxplot(width = .1, alpha = 0.9, fatten = NULL, show.legend = TRUE, legend='right') +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .1) + 
  theme(aspect.ratio = 1, axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.position = 'right')
d + annotate("text", x = 1:5, y= 1.2, label = brlabel1$Letter)
```



```{r}
#Total Root Length
rtlen <- ggboxplot(Root_Summary_1, x = "Treatment", y= "total root length (mm)", fill = "Treatment", palette = "jco", xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='wilcox.test', ref.group = "Control", hide.ns = TRUE) + theme(aspect.ratio = 1)
#Number of Branch Points
bp <- ggboxplot(Root_Summary_1, x = "Treatment", y= "number of branch points", fill = "Treatment", palette = "jco", xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)+ theme(aspect.ratio = 1)
# ggsave()
#Average Diameter
dia <- ggboxplot(Root_Summary_1, x = "Treatment", y= "average diameter (mm)", fill = "Treatment", palette = "jco", xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)+ theme(aspect.ratio = 1)
rtlen
bp
dia

#multi_plot_tips <- ggarrange(soil2021 + rremove("xlab"), soil_2022 + rremove("ylab"))
#multi_plot_tips

# create multi panel plot
multi_plot <- ggarrange(rtip + rremove("x.text"), rtlen + rremove("x.text"),  bp + rremove("x.text"),dia + rremove("x.text"), common.legend = TRUE, legend = 'top',labels = 'AUTO', rows=4, cols=1, font.label = list(size=20)) + theme(aspect.ratio =0.8)
multi_plot
ggarrange(rtip + rremove("x.text"), rtlen + rremove("x.text"), bp, dia, common.legend = TRUE, labels = 'AUTO')
#ggsave("Multipane root.tiff", plot = multi_plot)
```

```{r}

#Height
ggviolin(Root_Summary_1, x = "treatment", y= "height (mm)", fill = "treatment", palette = "jco", xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='anova', ref.group = "Control", hide.ns = TRUE)

#Volume
ggviolin(Root_Summary_1, x = "treatment", y= "volume (mm^3)", fill = "Treatment")
rtvol1 <- ggboxplot(Root_Summary_1, x = "Treatment", y= "volume (mm^3)", fill = "Treatment", palette = "jco", xlab = FALSE, width = 0.5)+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)
#Network Area
rtnet1 <- ggboxplot(Root_Summary_1, x = "Treatment", y= "network area (mm^2)", fill = "Treatment", palette = "jco", xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)
#Surface Area
rtsa1 <- ggboxplot(Root_Summary_1, x = "Treatment", y= "surface area (mm^2)", fill = "Treatment", palette = "jco", xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)

#combine everything together for a 4-panel figure
metaplot1 <- ggarrange(he1 + rremove("x.text"), rtvol1 + rremove("x.text"), rtnet1 + rremove("x.text"), rtsa1 + rremove("x.text"),  common.legend = TRUE, legend = 'bottom',labels = 'AUTO')
metaplot1

```
```{r}

root_sum_matrix1 <- Root_Summary_3[c(5:9,12:22)]

root_cor1 <- cor(root_sum_matrix1, use='complete.obs')

corrplot(root_cor1, type = 'full', order='hclust', hclust.method='ward.D2', tl.col='black', tl.srt=45)

corrplot(root_cor1, type = 'full', order='hclust', hclust.method='average', tl.col='black', tl.srt=45, addrect=4, method="circle")
res1 = cor.mtest(root_cor1, conf.level=0.99)
corrplot(root_cor1, type = 'full', order='hclust', hclust.method='average', tl.col='black', tl.srt=45, addrect=6, method="circle", p.mat = res1$p, sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.9, insig = 'label_sig', pch.col = 'grey20')
```

#Root tips
ggboxplot(Root_Summary, x = "Treatment", y= "Root tips", fill = "Treatment", palette = "jco", xlab = FALSE, width = 0.5)+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)
#Total Root Length
ggboxplot(Root_Summary, x = "Treatment", y= "Total root length (mm)", color = "Treatment", xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)
#Number of Branch Points
ggboxplot(Root_Summary, x = "Treatment", y= "Number of branch points", color = "Treatment", xlab = FALSE, width = 0.5, )+
  stat_compare_means(label = 'p.signif', method ='t.test', ref.group = "Control", hide.ns = TRUE)
#Average Diameter
ggboxplot(Root_Summary, x = "Treatment", y= "Average diameter (mm)", color = "Treatment", xlab = FALSE, width = 0.5,)+
  stat_compare_means(label = 'p.signif', method ='wilcox.test', ref.group = "Control", hide.ns = TRUE)
